Demo arXiv template

### 2022-11-22 paper/paper.qmd
more on scaling

@@

@@

Scaling is a specific variable transformation on the scale of the data. 
The purpose to treat scaling as a separate step from variable transformation is to make the step explicit. 
Scaling is usually in the form of $[x(s_i;t_j) - \alpha]/\gamma$. For example, $\alpha = \bar{x}(s; t)$ and $\gamma = \sigma(s; t)$ is used in a z-score standardisation to standardise the data by the its mean and standard deviation across all time and all space units and $\alpha = \min[x(s_i, t_j)]$ and $\gamma = \max[x(s_i, t_j)] - \min[x(s_i, t_j)]$ is used to standardise the variable into a unit interval.

@@

The purpose of normalising is for cross-comparison using the normal score.
This step can get criticism from analysts for ...

This step is usually in the end of the pipeline when the data have been processed into a univariate series. 
When the data are not distributed normally, a distributed is usually fitted before converting to the normal score via the normal reverse CDF function: $\Phi^{-1}(.)$: $\Phi^{-1}[F_{\eta}(x(\symbf{s}; \symbf{t}))]$, where $\eta \in H \subseteq \mathbb{R}^{d_{\eta}}$ is the distribution parameter and $d_{\eta}$ is the number of parameter of $\eta$. 
Sometimes, to account for the monthly variation in the data, the fitting is done separately for each month and the noramlising step can be written as $\Phi^{-1}[F_{\eta}^m(x(\symbf{s};t_{j^*}))]$, where $j^*$ is all the indexes that satisfy $j^* \mod 12 = m$ for each $m = 0, 1,\cdots, 11$.

@@

In public communication, indexes are often delivered in categorical grades, along with its underlying numerical values. The simplification step prescribes how the continuous index values are converted into the discrete grades. This process can be written with the piece-wise function:

>>>>

  - divide into pre and post-processing

@@

Variable transformation can be necessary for different reasons: 1) to change the shape of the variable, 2) to stablise the variance.

mention modelling? -> satisfy model assumption

@@

Sometimes, scaling, as a specific type of variable transformation, is merged into dimension reduction as a pre-step, i.e. transform the covariance matrix into correlation matrix before applying the Principal Component Analysis (PCA), but it worth to be separated into its own step to make it explicit.
Its main different from variable transformation is that it can be written in the form of $$[x(s_i;t_j) - \alpha]/\gamma.$$ 
For example, In a z-score standardisation, $\alpha = \bar{x}(s; t)$ and $\gamma = \sigma(s; t)$ are used to standardise the data by the its mean and standard deviation across all time and all space units
Another example is to scale the variable into a unit interval, which can be written as $\alpha = \min[x(s_i, t_j)]$ and $\gamma = \max[x(s_i, t_j)] - \min[x(s_i, t_j)]$.
While being a common technique in data pre-processing, scaling has an impact on the outliers of the data and analysts need to be cautious on checking its effect.

@@

The purpose of normalising is for cross-comparison using the normal score.
This step can get criticism from analysts for being unnecessarily forcing the scale to the normal scale.
This step is usually in the end of the pipeline when the data have been processed into a univariate series. 
When the data are not distributed normally, a distributed is usually fitted before converting to the normal score via the normal reverse CDF function: $\Phi^{-1}(.)$: $\Phi^{-1}[F_{\eta}(x(\symbf{s}; \symbf{t}))]$, where $\eta \in H \subseteq \mathbb{R}^{d_{\eta}}$ is the distribution parameter and $d_{\eta}$ is the number of parameter of $\eta$. 
Sometimes, to account for the monthly variation in the data, the fitting is done separately for each month and the noramlising step can be written as $\Phi^{-1}[F_{\eta}^m(x(\symbf{s};t_{j^*}))]$, where $j^*$ is all the indexes that satisfy $j^* \mod 12 = m$ for each $m = 0, 1,\cdots, 11$.

@@

In public communication, indexes are often delivered in categorical grades, along with its underlying numerical values. The simplification step, sometimes can also be called discretisation, prescribes how the continuous index values are converted into the discrete grades. This process can be written with the piece-wise function:

### 2022-11-22 paper/paper.tex
more on scaling

\usepackage{arxiv}
\usepackage{orcidlink}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\usepackage{fontspec}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Demo arXiv template},
  pdfauthor={H. Sherry Zhang; Collaborators},
  pdfkeywords={indexes, data pipeline, software design},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, sharp corners, breakable, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

@@

This is what rendered by \texttt{\textbackslash{}symbf\{x\}}:
\(uvwxyz, \boldsymbol{uvwxyz}, \pmb{uvwxyz}, \symbf{uvwxyz}\)

@@

@@

Scaling is a specific variable transformation on the scale of the data.
The purpose to treat scaling as a separate step from variable
transformation is to make the step explicit. Scaling is usually in the
form of \([x(s_i;t_j) - \alpha]/\gamma\). For example,
\(\alpha = \bar{x}(s; t)\) and \(\gamma = \sigma(s; t)\) is used in a
z-score standardisation to standardise the data by the its mean and
standard deviation across all time and all space units and
\(\alpha = \min[x(s_i, t_j)]\) and
\(\gamma = \max[x(s_i, t_j)] - \min[x(s_i, t_j)]\) is used to
standardise the variable into a unit interval.

@@

The purpose of normalising is for cross-comparison using the normal
score. This step can get criticism from analysts for \ldots{}

This step is usually in the end of the pipeline when the data have been
processed into a univariate series. When the data are not distributed
normally, a distributed is usually fitted before converting to the
normal score via the normal reverse CDF function: \(\Phi^{-1}(.)\):
\(\Phi^{-1}[F_{\eta}(x(\symbf{s}; \symbf{t}))]\), where
\(\eta \in H \subseteq \mathbb{R}^{d_{\eta}}\) is the distribution
parameter and \(d_{\eta}\) is the number of parameter of \(\eta\).
Sometimes, to account for the monthly variation in the data, the fitting
is done separately for each month and the noramlising step can be
written as \(\Phi^{-1}[F_{\eta}^m(x(\symbf{s};t_{j^*}))]\), where
\(j^*\) is all the indexes that satisfy \(j^* \mod 12 = m\) for each
\(m = 0, 1,\cdots, 11\).

>>>>

\usepackage{arxiv}
\usepackage{orcidlink}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Demo arXiv template},
  pdfauthor={H. Sherry Zhang; Collaborators},
  pdfkeywords={indexes, data pipeline, software design},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden]}{\end{tcolorbox}}\fi

@@

\begin{itemize}
\tightlist
\item
  divide into pre and post-processing
\end{itemize}

@@

@@

Variable transformation can be necessary for different reasons: 1) to
change the shape of the variable, 2) to stablise the variance.

mention modelling? -\textgreater{} satisfy model assumption

@@

Sometimes, scaling, as a specific type of variable transformation, is
merged into dimension reduction as a pre-step, i.e.~transform the
covariance matrix into correlation matrix before applying the Principal
Component Analysis (PCA), but it worth to be separated into its own step
to make it explicit. Its main different from variable transformation is
that it can be written in the form of \[[x(s_i;t_j) - \alpha]/\gamma.\]
For example, In a z-score standardisation, \(\alpha = \bar{x}(s; t)\)
and \(\gamma = \sigma(s; t)\) are used to standardise the data by the
its mean and standard deviation across all time and all space units
Another example is to scale the variable into a unit interval, which can
be written as \(\alpha = \min[x(s_i, t_j)]\) and
\(\gamma = \max[x(s_i, t_j)] - \min[x(s_i, t_j)]\). While being a common
technique in data pre-processing, scaling has an impact on the outliers
of the data and analysts need to be cautious on checking its effect.

@@

The purpose of normalising is for cross-comparison using the normal
score. This step can get criticism from analysts for being unnecessarily
forcing the scale to the normal scale. This step is usually in the end
of the pipeline when the data have been processed into a univariate
series. When the data are not distributed normally, a distributed is
usually fitted before converting to the normal score via the normal
reverse CDF function: \(\Phi^{-1}(.)\):
\(\Phi^{-1}[F_{\eta}(x(\symbf{s}; \symbf{t}))]\), where
\(\eta \in H \subseteq \mathbb{R}^{d_{\eta}}\) is the distribution
parameter and \(d_{\eta}\) is the number of parameter of \(\eta\).
Sometimes, to account for the monthly variation in the data, the fitting
is done separately for each month and the noramlising step can be
written as \(\Phi^{-1}[F_{\eta}^m(x(\symbf{s};t_{j^*}))]\), where
\(j^*\) is all the indexes that satisfy \(j^* \mod 12 = m\) for each
\(m = 0, 1,\cdots, 11\).
